Universal Anonymous Signatures

### 2022-06-27 prism-backend/research/gsac/formal-uas.tex
Added a paragraph with real-world sample use cases. Fixed some todos.

Next we formalize our notion of Universal Anonymous Signatures
(\UAS). We define its syntax, model, and give a generic construction which
we prove secure according to our model.

@@

\begin{description}
\item[$\fissue: (\usk,\scred,\attrs)
  \rightarrow 0/1$.] Chosen by each issuer within a family of functions
  \famfissue, the issuance function defines the customized conditions required
  by the issuer to grant a credential over attributes \attrs, when requested by
  a user with secret key \usk, potentially leveraging a set of previously
  obtained credentials \scred. \fissue returns $1$ to accept a request, $0$ to
  reject it.
\item[$\feval: (\usk,\scred,\msg)
  \rightarrow (\yeval^0,\yeval^1)$.] Signing evaluation functions, from a
  family of functions \famfeval, can be set on a per-signature basis. They
  receive the user secret key \usk, credentials \scred, and message to be signed
  \msg. \feval outputs two values, $\yeval^0$ and $\yeval^1$, where the former
  is revealed in plaintext along with the signature, and $\yeval^1$ is
  encrypted. Both can be used to modulate the bevahiour of \finsp, and must
  belong in a well defined set \rngfeval. For readability, we write $\Yeval =
  (\yeval^0,\yeval^1)$.
\item[$\finsp: (\Yeval,\usk,\scred,\msg) \rightarrow \yinsp$.]
  Chosen by openers from a family of functions \famfinsp. The opening
  functions define what utility value, derived from the user's secret key,
  credentials, and signed message, should be extractable by an opener. Its
  behaviour can be modulated by \Yeval, and its output a value \yinsp, which
  must belong in a well defined set \rngfinsp.
\end{description}

\feval and \finsp may seem redundant. However, observe that the former is chosen
on a per-signature basis, and probably after negotiation between signer and
verifier. As such, it governs the information revealed at signing time. On the
other hand, \finsp governs the information that may be revealed after signing.
Moreover, it is defined by openers, and even though signers and verifiers
may negotiate to use any opener for a given signature, it is not possible for
an opener to change the chosen function.

\begin{definition}{(\CUASGen restrictions)}
  \label{def:uas-restrictions}
  Let \CUASGen be a generic construction of an \UAS scheme, secure according to
  the model defined in \secref{ssec:model-uas}, and three functions $\fissue^a$,
  $\feval^b$, $\finsp^c$ compatible with the previous definitions. We say that
  \CUASGen, instantiated with $\fissue^a$, $\feval^b$, and $\finsp^c$, is a
  $(\fissue^a,\feval^b,\finsp^c)$-restriction of \CUASGen.
\end{definition}

To summarize the previous, we depict in \figref{fig:uas-signature} the structure
of an \UAS signature, including its main components and how they enable the
previously stated generalization claims.

\begin{figure}[ht!]
  \centering
  %\input{figures/uas-signature.tex}
  \caption{Structure of an \UAS signature.}
  \label{fig:uas-signature}
\end{figure}

@@

We define the security expected from an \UAS scheme using a game-based approach.
An \UAS scheme must ensure privacy at credential issuance, meaning that two
runs of the credential issuance protocol should not leak any information beyond
necessary about the user requesting the credential. Note that this is now
crucial for privacy since, as AC schemes, \UAS allows multiple credentials per
user -- thus, knowing what credential a user requests can lead to user
profiling. It must also ensure signature privacy, meaning that no information
can be learned, from a signature, by third parties beyond what the user agrees
to reveal, and what may be extractable by the chosen opener.
Issuance unforgeability requires that \UAS schemes ensure that the claims used
to support a credential request cannot be falsified (issuance unforgeability).
Similarly, signature unforgeability requires that no adversary should be able to
falsify the correctness proofs for the claims included in the signatures. In
both cases (issuance and signature unforgeability), we introduce the notion of
traceability in GS, requiring that all credentials involved in the corresponding
process have to be associated to a known user. Finally, \UAS schemes inherit the
useful non-frameability property of GS schemes which, even in the worst-case
scenario of having a corrupt issuer. In this case, although we cannot ensure
anything regarding correctness of the claims, we can ensure that no honest user
can be framed with having created a signature that s/he did not create --
although the concept may get a bit blurred, as there is no traditional opening.
%
We describe next the global variables and oracles that we use in our modelling.

@@

        \procedure{$\OBTISS(\cid,\uid,\iid,\attrs,\scid)$}{%
          \pcif \uid \notin \HU: \pcreturn \bot \\
          \pcif \iid \notin \HI: \pcreturn \bot \\
          \pcif \CRED[\cid] \neq \bot \lor \CCRED[\cid] \neq \bot: \pcreturn \bot \\
          \langle \cred, \utrans \rangle \gets
          \langle \Obtain(\UK[\uid],\CRED[\scid],\attrs), \\
          \hspace*{60pt} \Issue(\PRVIK[\iid],\ISR[\scid],\attrs)
          \rangle \\
          \trans[\cid] \gets \utrans \\
          \CRED[\cid] \gets (\uid, \cred, \iid, \attrs, \scid, \siid) \\
          \pcreturn \top \\
        }

@@

        \procedure{$\SIGN(\oid,\uid,\scid,\msg,\feval)$}{%
          \pcif \uid \notin \HU \cup \lbrace \bot \rbrace: \pcreturn \bot \\
          \pcif \uid = \bot \land \exists \cid' \in \scid~\suchthat~
          \cid' \notin \CCRED: \\
          \pcind \pcreturn \bot \\
          \pcif \uid = \bot: \uid \gets \OWNR[\scid] \\
          \pcif \feval \notin \famfeval: \pcreturn \bot \\
          \Sig \gets \Sign(\UK[\uid],\PUBOK[\oid],\CRED[\scid],\msg,\feval) \\
          \Yeval=(\yeval^0,\yeval^1) \gets \feval(\UK[\uid],\CRED[\scid],\msg) \\
          \pcif \Yeval \notin \rngfeval: \pcreturn \bot \\
          \SIG[\uid] \gets \SIG[\uid] \cup
          \lbrace (\oid,\scid,\Sig,\Yeval,\msg,\feval) \rbrace \\
          \pcreturn \Sig \\
        }

@@

        \procedure{$\CHALb(\oid,\cuid_{0,1},\cscid_{0,1},\msg,\feval)$}{%
          \pcif \cuid_0 \notin \HU \lor \cuid_1 \notin \HU: \pcreturn \bot \\
          \pcif \oid \in \CO: \pcreturn \bot \\
          \pcif \feval \notin \famfeval: \pcreturn \bot \\          
          \Yeval = (\yeval^0,\yeval^1) \gets \feval(\UK[\cuid_0],\CRED[\cscid_0],
          \msg) \\
          \TYeval = (\tyeval^0,\tyeval^1) \gets
          \feval(\UK[\cuid_1],\CRED[\cscid_1],\msg) \\
          \pcif \yeval^0 \neq \tyeval^0: \pcreturn \bot \\
          \pcif \Yeval \notin \rngfeval \lor \TYeval \notin \rngfeval:
          \pcreturn \bot \\
          \pcif \PUBIK[\cscid_0] \neq \PUBIK[\cscid_1]: \pcreturn \bot \\
          \cSig_b \gets \Sign(\UK[\cuid_b],\PUBOK[\oid], \\
          \hspace*{71pt}\CRED[\cscid_b],\msg,\feval) \\
          \cSig_{1-b} \gets \Sign(\UK[\cuid_{1-b}],\PUBOK[\oid], \\
          \hspace*{80pt}\CRED[\cscid_{1-b}],\msg,\feval) \\          
          \CSIG[\cSig_b] \gets 
          \lbrace (\oid,\cuid_b,\cscid_b,\Yeval,\msg,\feval,\\
          \hspace*{74pt}\cuid_{1-b},\cSig_{1-b},\cscid_{1-b},\TYeval)\rbrace \\
          \pcreturn \cSig_b
        }

@@

The issuance anonymity game captures that, given polymomially many
credential issuance interactions, the adversary gains no information about the
user(s) behind them. For this, we allow the adversary to control targetted
issuers and openers (although it may also add honest ones), and add honest and
corrupt users. The adversary can also obtain credentials, produce signatures,
and open them. We need to make some limitations to prevent trivial wins: first,
when querying the oracle for obtaining a challenge credential out of two
possible user-credential set pairs, the issuance protocol must either succeed or
fail for both; also, the adversary can open signatures produced with credential
sets containing challenge credentials $\cred_b$, only if the output of \Open is
the same as with the complementary $\cred_{1-b}$ 
\todo{this is not done in the oracles. Fix.} (note that any other signature
can be open).
Intuitively, if the adversary cannot distinguish arbitrary runs with all oracles
(provided the previous constraint) when he gets challenge credentials $\cred_b$
from runs with credentials $\cred_{1-b}$, then the issuance protocol is
privacy-preserving.

@@

\begin{figure}[htp!]
    \procedure[linenumbering]{$\ExpForgeIssue(1^\secpar)$}{%
      \parm \gets \Setup(1^\secpar) \\
      \cid \gets \adv^{\Oforgeissue}(\parm) \\
      \pcif \trans[\cid] = \bot \lor \CRED[\cid] = \bot: \pcreturn 0 \\
      \textrm{Parse}~\CRED[\cid]~\textrm{as}~(\cdot,\cdot,\iid,\cdot,\cdot);~
      \IK[\iid]~\textrm{as}~((\ipk,\fissue),\cdot) \\
      (\usk,\scred,\attrs_{\scred}) \gets \ExtractIssue(\trans[\cid]) \\
      \pcif \fissue(\usk,\scred,\ATTR[\cid]) = 0 \lor
      \exists \cred \in \scred~\suchthat~\IdentifyCred(\usk,\attrs_{\cred},\cred) = 0: \\
      \pcind \pcreturn 1 \\
      \pcif \scred \neq \emptyset \land
      \nexists \uid~\suchthat~\IdentifyUK(\uid,\usk) = 1: \pcreturn 1~
      \pccomment{\todo{New! Update proofs.}} \\
      \pcreturn 0
    }
  \caption{Experiment for unforgeability of credential issuance in \UAS schemes.}
  \label{fig:exp-uas-unfor-issue}
\end{figure}

@@

\begin{figure}[htp!]
  \procedure[linenumbering]{$\ExpNonframe(1^\secpar)$}{%
    \parm \gets \Setup(1^\secpar) \\
    (\oid,\siid,\Sig=(\sig,\yeval),\msg,\feval,\yinsp,\iproof) \gets
    \adv^{\Oframe}(\parm) \\
    \pcif \Verify(\PUBOK[\oid],\PUBIK[\siid],\Sig,\msg,\feval) = 0:
    \pcreturn 0 \\
    \pcif \Judge(\PUBOK[\oid],\PUBIK[\siid],\yinsp,\iproof,\Sig,\msg) = 0:
    \pcreturn 0 \\
    (\usk,\scred,\attrs_{\scred},\tyeval^1,\yinsp') \gets
    \ExtractSign(\oid,\siid,\Sig,\msg,\feval) \\
    \Yeval=(\yeval^0,\yeval^1) \gets \feval(\usk,\scred,\msg)
    \pccomment{\todo{Added this line -- check line refs.}} \\
    \pcif \yeval \neq \yeval^0 \lor \tyeval^1 \neq \yeval^1
    \lor \Yeval \notin \rngfeval: \pcreturn 1 \\
    \pcif \finsp(\Yeval,\usk,\scred,\msg) \neq \yinsp
    \lor \yinsp \neq \yinsp': \pcreturn 1 \\    
    \pcif \nexists \uid \in \HU~\suchthat~\UK[\uid] = \usk: \pcreturn 0 \\
    \pcif \nexists (\cdot,\cdot,\Sig,\Yeval,\msg,\feval)
    \in \SIG[\uid]: \pcreturn 1 \\
    \pcreturn 0
  }
  \caption{Experiment for non-frameability on \UAS schemes.}
  \label{fig:exp-uas-frame}
\end{figure}

>>>>

Next we formalize our notion of Universal Anonymous Signatures
(\UAS). We define its syntax, model, and give a generic construction, which
we prove secure according to our model.

@@

\begin{description}
\item[$\fissue: (\usk,\scred,\attrs)
  \rightarrow 0/1$.] Chosen by each issuer within a family of functions
  \famfissue, the issuance function defines the customized conditions required
  by the issuer to grant a credential over attributes \attrs, when requested by
  a user with secret key \usk, potentially leveraging a set of previously
  obtained credentials \scred. \fissue returns $1$ to accept a request, $0$ to
  reject it.
\item[$\feval: (\usk,\scred,\msg)
  \rightarrow (\yeval^0,\yeval^1)$.] Signing evaluation functions, from a
  family of functions \famfeval, can be set on a per-signature basis. They
  receive the user secret key \usk, credentials \scred, and message to be signed
  \msg. \feval outputs two values, $\yeval^0$ and $\yeval^1$, where the former
  is revealed in plaintext along with the signature, and $\yeval^1$ is
  encrypted. Both can be used to modulate the bevahiour of \finsp, and must
  belong in a well defined set \rngfeval. For readability, we write $\Yeval =
  (\yeval^0,\yeval^1)$.
\item[$\finsp: (\Yeval,\usk,\scred,\msg) \rightarrow \yinsp$.]
  Chosen by openers from a family of functions \famfinsp. The opening
  functions define what utility value, derived from the user's secret key,
  credentials used for signing, and signed message, should be extractable by an
  opener. Its behaviour can be modulated by \Yeval, and it outputs a value
  \yinsp, which must belong in a well defined set \rngfinsp.
\end{description}

\feval and \finsp may seem redundant. However, observe that the former is chosen
on a per-signature basis, probably after negotiation between signer and
verifier. As such, it governs the information revealed at signing time. On the
other hand, \finsp governs the information that may be revealed after signing.
Moreover, it is defined by openers, and even though signers and verifiers
may negotiate to use any opener for a given signature, it is not possible for
an opener to change the chosen function.

\begin{definition}{(\CUASGen restrictions)}
  \label{def:uas-restrictions}
  Let \CUASGen be a construction of an \UAS scheme, secure according to
  the model defined in \secref{ssec:model-uas}, and three functions $\fissue^a$,
  $\feval^b$, $\finsp^c$ compatible with the previous definitions. We say that
  \CUASGen, instantiated with $\fissue^a$, $\feval^b$, and $\finsp^c$, is a
  $(\fissue^a,\feval^b,\finsp^c)$-restriction of \CUASGen.
\end{definition}

% To summarize the previous, we depict in \figref{fig:uas-signature} the structure
% of an \UAS signature, including its main components and how they enable the
% previously stated generalization claims.

% \begin{figure}[ht!]
%   \centering
%   %\input{figures/uas-signature.tex}
%   \caption{Structure of an \UAS signature.}
%   \label{fig:uas-signature}
% \end{figure}

@@

We define the security expected from an \UAS scheme using a game-based approach.
An \UAS scheme must ensure privacy at credential issuance, meaning that two
runs of the credential issuance protocol should not leak any information beyond
necessary about the user requesting the credential. Note that this is now
crucial for privacy since, as AC schemes, \UAS allows multiple credentials per
user -- thus, knowing what user is requesting which credential can lead to user
profiling. It must also ensure signature privacy, meaning that no information
can be learned by third parties from a signature, beyond what the user chooses
to reveal, and what may be extractable by the chosen opener. Issuance
unforgeability requires that \UAS schemes ensure that the claims used to support
a credential request cannot be falsified (issuance unforgeability). Similarly,
signature unforgeability requires that no adversary should be able to falsify
the correctness proofs for the claims included in the signatures. In both cases
(issuance and signature unforgeability), we require all credentials involved in
the corresponding process to be associated to a known user. Finally, \UAS
schemes inherit the useful non-frameability property of GS schemes where even
the issuer is corrupt. In this case, we can ensure that no honest user
can be framed with having created a signature that s/he did not create --
although the concept may get a bit blurred, as there is no traditional opening.
%
We describe next the global variables and oracles that we use in our modelling.

@@

        \procedure{$\OBTISS(\cid,\uid,\iid,\attrs,\scid)$}{%
          \pcif \uid \notin \HU: \pcreturn \bot \\
          \pcif \iid \notin \HI: \pcreturn \bot \\
          \pcif \exists \cid' \in \scid~\suchthat~\cid' \in \CCRED: \pcreturn \bot \\
          \pcif \CRED[\cid] \neq \bot \lor \CCRED[\cid] \neq \bot: \pcreturn \bot \\
          \langle \cred, \utrans \rangle \gets
          \langle \Obtain(\UK[\uid],\CRED[\scid],\attrs), \\
          \hspace*{60pt} \Issue(\PRVIK[\iid],\ISR[\scid],\attrs)
          \rangle \\
          \trans[\cid] \gets \utrans \\
          \CRED[\cid] \gets (\uid, \cred, \iid, \attrs, \scid, \siid) \\
          \pcreturn \top \\
        }

@@

        \procedure{$\SIGN(\uid,\oid,\scid,\msg,\feval)$}{%
          \pcif \uid \notin \HU: \pcreturn \bot \\
          \pcif \exists \cid,\cid' \in \scid~\suchthat~\cid \in \CRED \land
          \cid' \in \CCRED: \\
          \pcind \pcreturn \bot \\
          \pcif \feval \notin \famfeval: \pcreturn \bot \\
          \Sig \gets \Sign(\UK[\uid],\PUBOK[\oid],\CRED[\scid],\msg,\feval) \\
          \Yeval=(\yeval^0,\yeval^1) \gets \feval(\UK[\uid],\CRED[\scid],\msg) \\
          \pcif \Yeval \notin \rngfeval: \pcreturn \bot \\
          \SIG[\uid] \gets \SIG[\uid] \cup
          \lbrace (\oid,\scid,\Sig,\Yeval,\msg,\feval) \rbrace \\
          \pcreturn \Sig \\
        }

@@

        \procedure{$\CHALb(\oid,\cscid_{0,1},\msg,\feval)$}{%
          \pcif \uid \notin \HU: \pcreturn \bot \\
          \pcif \pcfor d \in \bin~ \exists \cid,\cid' \in \scid_d~\suchthat \\
          \pcind \cid \in \CRED \land \cid' \in \CCRED: \pcreturn \bot \\
          \pcif \oid \in \CO: \pcreturn \bot \\
          \pcif \feval \notin \famfeval: \pcreturn \bot \\          
          \Yeval = (\yeval^0,\yeval^1) \gets \feval(\UK[\cuid_0],\CRED[\cscid_0],
          \msg) \\
          \TYeval = (\tyeval^0,\tyeval^1) \gets
          \feval(\UK[\cuid_1],\CRED[\cscid_1],\msg) \\
          \pcif \yeval^0 \neq \tyeval^0: \pcreturn \bot \\
          \pcif \Yeval \notin \rngfeval \lor \TYeval \notin \rngfeval:
          \pcreturn \bot \\
          \pcif \PUBIK[\cscid_0] \neq \PUBIK[\cscid_1]: \pcreturn \bot \\
          \cSig_b \gets \Sign(\UK[\cuid_b],\PUBOK[\oid], \\
          \hspace*{71pt}\CRED[\cscid_b],\msg,\feval) \\
          \cSig_{1-b} \gets \Sign(\UK[\cuid_{1-b}],\PUBOK[\oid], \\
          \hspace*{80pt}\CRED[\cscid_{1-b}],\msg,\feval) \\          
          \CSIG[\cSig_b] \gets 
          \lbrace (\oid,\cuid_b,\cscid_b,\Yeval,\msg,\feval,\\
          \hspace*{74pt}\cuid_{1-b},\cSig_{1-b},\cscid_{1-b},\TYeval)\rbrace \\
          \pcreturn \cSig_b
        }

@@

The issuance anonymity game captures that, given polymomially many
credential issuance interactions, the adversary gains no information about the
user(s) behind them. For this, we allow the adversary to control targetted
issuers and openers (although it may also add honest ones), and add honest and
corrupt users. The adversary can also obtain credentials, produce signatures,
and open them. We need to make some limitations to prevent trivial wins: first,
when querying the oracle for obtaining a challenge credential out of two
possible user-credential set pairs, the issuance protocol must either succeed or
fail for both; also, the adversary can open signatures produced with credential
sets containing challenge credentials $\cred_b$, only if the output of \Open is
the same as with the complementary $\cred_{1-b}$ (note that any other signature
can be open).
Intuitively, if the adversary cannot distinguish arbitrary runs with all oracles
(provided the previous constraint) when he gets challenge credentials $\cred_b$
from runs with credentials $\cred_{1-b}$, then the issuance protocol is
privacy-preserving.

@@

\begin{figure}[htp!]
    \procedure[linenumbering]{$\ExpForgeIssue(1^\secpar)$}{%
      \parm \gets \Setup(1^\secpar) \\
      \cid \gets \adv^{\Oforgeissue}(\parm) \\
      \pcif \trans[\cid] = \bot \lor \CRED[\cid] = \bot: \pcreturn 0 \\
      \textrm{Parse}~\CRED[\cid]~\textrm{as}~(\cdot,\cdot,\iid,\cdot,\cdot);~
      \IK[\iid]~\textrm{as}~((\ipk,\fissue),\cdot) \\
      (\usk,\scred,\attrs_{\scred}) \gets \ExtractIssue(\trans[\cid]) \\
      \pcif \fissue(\usk,\scred,\ATTR[\cid]) = 0 \lor
      \exists \cred \in \scred~\suchthat~\IdentifyCred(\usk,\attrs_{\cred},\cred) = 0: \\
      \pcind \pcreturn 1 \\
      \pcreturn 0
    }
  \caption{Experiment for unforgeability of credential issuance in \UAS schemes.}
  \label{fig:exp-uas-unfor-issue}
\end{figure}

@@

\begin{figure}[htp!]
  \procedure[linenumbering]{$\ExpNonframe(1^\secpar)$}{%
    \parm \gets \Setup(1^\secpar) \\
    (\oid,\siid,\Sig=(\sig,\yeval),\msg,\feval,\yinsp,\iproof) \gets
    \adv^{\Oframe}(\parm) \\
    \pcif \Verify(\PUBOK[\oid],\PUBIK[\siid],\Sig,\msg,\feval) = 0:
    \pcreturn 0 \\
    \pcif \Judge(\PUBOK[\oid],\PUBIK[\siid],\yinsp,\iproof,\Sig,\msg) = 0:
    \pcreturn 0 \\
    (\usk,\scred,\attrs_{\scred},\tyeval^1,\yinsp') \gets
    \ExtractSign(\oid,\siid,\Sig,\msg,\feval) \\
    \Yeval=(\yeval^0,\yeval^1) \gets \feval(\usk,\scred,\msg) \\
    \pcif \yeval \neq \yeval^0 \lor \tyeval^1 \neq \yeval^1
    \lor \Yeval \notin \rngfeval: \pcreturn 1 \\
    \pcif \finsp(\Yeval,\usk,\scred,\msg) \neq \yinsp
    \lor \yinsp \neq \yinsp': \pcreturn 1 \\    
    \pcif \nexists \uid \in \HU~\suchthat~\UK[\uid] = \usk: \pcreturn 0 \\
    \pcif \nexists (\cdot,\cdot,\Sig,\Yeval,\msg,\feval)
    \in \SIG[\uid]: \pcreturn 1 \\
    \pcreturn 0
  }
  \caption{Experiment for non-frameability on \UAS schemes.}
  \label{fig:exp-uas-frame}
\end{figure}
